% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[BoldFont = SF Pro Rounded Semibold]{SF Pro Rounded}
  \setmathfont[]{STIX Two Math}
\fi
\usefonttheme{serif} % use mainfont rather than sansfont for slide text
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={305 Lecture 1.3 - Validity},
  pdfauthor={Brian Weatherson},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\let\Tiny=\tiny

 \setbeamertemplate{navigation symbols}{} 

% \usetheme{Madrid}
 \usetheme[numbering=none, progressbar=foot]{metropolis}
 \usecolortheme{wolverine}
 \usepackage{color}
 \usepackage{MnSymbol}
% \usepackage{movie15}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\DeclareSymbolFont{symbolsC}{U}{txsyc}{m}{n}
\DeclareMathSymbol{\boxright}{\mathrel}{symbolsC}{128}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\usepackage{tikz-qtree}
% \usepackage{markdown}
%\RequirePackage{bussproofs}
\usetikzlibrary{arrows.meta}
\RequirePackage[tableaux]{prooftrees}
\forestset{line numbering, close with = x}
% Allow for easy commas inside trees
\renewcommand{\,}{\text{, }}


\usepackage{tabulary}

\usepackage{open-logic-config}

\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\AtBeginSection[]
{
\begin{frame}
	\Huge{\color{darkblue} \insertsection}
\end{frame}
}

\renewenvironment*{quote}	
	{\list{}{\rightmargin   \leftmargin} \item } 	
	{\endlist }

\definecolor{darkgreen}{rgb}{0,0.7,0}
\definecolor{darkblue}{rgb}{0,0,0.8}

\newcommand{\starttab}{\begin{center}
\vspace{6pt}
\begin{tabular}}

\newcommand{\stoptab}{\end{tabular}
\vspace{6pt}
\end{center}
\noindent}


\newcommand{\sif}{\rightarrow}
\newcommand{\siff}{\leftrightarrow}
\newcommand{\EF}{\end{frame}}


\newcommand{\TreeStart}[1]{
%\end{frame}
\begin{frame}
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=north}}
%\Tree
}

\newcommand{\TreeEnd}{
\end{tikzpicture}
%\end{center}
}

\newcommand{\DisplayArg}[2]{
\begin{enumerate}
{#1}
\end{enumerate}
\vspace{-6pt}
\hrulefill

%\hspace{14pt} #2
%{\addtolength{\leftskip}{14pt} #2}
\begin{quote}
{\normalfont #2}
\end{quote}
\vspace{12pt}
}

\newenvironment{ProofTree}[1][1]{
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=south}}
}
{
\end{tikzpicture}
\end{center}
}

\newcommand{\TreeFrame}[2]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\begin{prooftree}{}
#1
\end{prooftree}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#2
%\end{markdown}
\end{columns}
}

\newcommand{\ScaledTreeFrame}[3]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\scalebox{#1}{
\begin{prooftree}{}
#2
\end{prooftree}
}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#3
%\end{markdown}
\end{columns}
}

\usepackage[bb=boondox]{mathalfa}
\DeclareMathAlphabet{\mathbx}{U}{BOONDOX-ds}{m}{n}
\SetMathAlphabet{\mathbx}{bold}{U}{BOONDOX-ds}{b}{n}
\DeclareMathAlphabet{\mathbbx} {U}{BOONDOX-ds}{b}{n}


\newenvironment{oltableau}{\center\tableau{}} %wff format={anchor = base west}}}
       {\endtableau\endcenter}
       
\newcommand{\formula}[1]{$#1$}

\usepackage{tabulary}
\usepackage{booktabs}

\def\begincols{\begin{columns}}
\def\begincol{\begin{column}}
\def\endcol{\end{column}}
\def\endcols{\end{columns}}

\usepackage[italic]{mathastext}
\usepackage{nicefrac}

\definecolor{mygreen}{RGB}{0, 100, 0}
\definecolor{mypink2}{RGB}{219, 48, 122}
\definecolor{dodgerblue}{RGB}{30,144,255}

%\def\True{\textcolor{dodgerblue}{\text{T}}}
%\def\False{\textcolor{red}{\text{F}}}

\def\True{\mathbb{T}}
\def\False{\mathbb{F}}

% This is because arguments didn't have enough space after them
\usepackage{etoolbox}
\AfterEndEnvironment{description}{\vspace{9pt}}
\AfterEndEnvironment{oltableau}{\vspace{9pt}}
\BeforeBeginEnvironment{oltableau}{\vspace{9pt}}
\AfterEndEnvironment{center}{\vspace{12pt}}
\BeforeBeginEnvironment{tabular}{\vspace{9pt}}

\setlength\heavyrulewidth{0pt}
\setlength\lightrulewidth{0pt}

%\def\toprule{}
%\def\bottomrule{}
%\def\midrule{}

\setbeamertemplate{caption}{\raggedright\insertcaption}

\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{305 Lecture 1.3 - Validity}
\author{Brian Weatherson}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{Plan for This Lecture}
\protect\hypertarget{plan-for-this-lecture}{}
We're discussing the nature of validity.
\end{frame}

\begin{frame}{Associated Reading}
\protect\hypertarget{associated-reading}{}
forall x, chapter 2, ``The Scope of Logic''.
\end{frame}

\begin{frame}{Virtues of Arguments}
\protect\hypertarget{virtues-of-arguments}{}
What are some virtues of arguments that you can think of? What makes for
a good argument?
\end{frame}

\begin{frame}{Virtues of Arguments}
\protect\hypertarget{virtues-of-arguments-1}{}
\begin{itemize}[<+->]
\tightlist
\item
  Premises are true.
\item
  Premises are known.
\item
  Premises are accepted in debate.
\item
  Premises are evidence for the conclusion.
\item
  Premises guarantee the truth of the conclusion.
\end{itemize}
\end{frame}

\begin{frame}{Logic}
\protect\hypertarget{logic}{}
To a first approximation, logic is about the last of these virtues.

\begin{itemize}
\tightlist
\item
  Does the truth of a set of premises \textbf{guarantee} the truth of
  the conclusion? \pause
\item
  If so, we'll say the argument is \textbf{valid}.
\end{itemize}
\end{frame}

\begin{frame}{Guarantee}
\protect\hypertarget{guarantee}{}
What is it for the premises to guarantee truth?

\begin{itemize}
\tightlist
\item
  It's for every \textbf{case} where the premises are true to also be a
  case where the conclusion is true. \pause
\item
  That's nice, you might say, but what's a case?
\end{itemize}
\end{frame}

\begin{frame}{Cases}
\protect\hypertarget{cases}{}
\begin{itemize}
\tightlist
\item
  This, I'm afraid, we're going to leave a little abstract.
\item
  As the book in effect notes, it's probably best to leave this as a
  placeholder.
\item
  Different notions of what a case are get you different notions of what
  arguments are valid.
\item
  But you get into deep metaphysical waters here pretty fast; let's
  retreat to some defintions.
\end{itemize}
\end{frame}

\begin{frame}{Terminology}
\protect\hypertarget{terminology}{}
When an argument, with premises \(P_1, P_2, \dots, P_n\) and conclusion
\(C\) is valid, we'll say all the following things are true. \pause

\begin{itemize}
\tightlist
\item
  The premises \textbf{entail} the conclusion. \pause
\item
  The conclusion \textbf{follows from} the premises. \pause
\item
  The conclusion is a \textbf{consequence} of the premises.
\end{itemize}
\end{frame}

\begin{frame}{Invalid}
\protect\hypertarget{invalid}{}
\begin{itemize}
\tightlist
\item
  An argument is \textbf{invalid} just in case it is not valid. \pause
\item
  That is, it is \textbf{invalid} if there is a case where the premises
  are true and the conclusion false.
\end{itemize}
\end{frame}

\begin{frame}{Structure}
\protect\hypertarget{structure}{}
\begin{itemize}
\tightlist
\item
  Logic is largely about validity, about which premises guarantee the
  truth of which conclusions.
\item
  It's about a special kind of guarantee, one that holds in virtue of
  the \textbf{structure} or \textbf{form} of the premises and
  conclusion.
\end{itemize}
\end{frame}

\begin{frame}{A Valid Argument}
\protect\hypertarget{a-valid-argument}{}
\begin{description}
\tightlist
\item[~]
All kangaroos are wise.
\item[~]
Skippy is a kangaroo.
\item[\(\therefore\)]
Skippy is wise.
\end{description}
\end{frame}

\begin{frame}{Non-Structural Guarantee}
\protect\hypertarget{non-structural-guarantee}{}
Compare the Skippy argument with this one.

\begin{description}
\tightlist
\item[~]
Today is Wednesday.
\item[\(\therefore\)]
Tomorrow is Thursday.
\end{description}

The premise guarantees the truth of the conclusion, but it does so in
virtue of meanings of `Wednesday' and `Thursday', not structural
features.
\end{frame}

\begin{frame}{Form and Structure}
\protect\hypertarget{form-and-structure}{}
\begin{itemize}
\tightlist
\item
  In the book they talk about the \textbf{form} of arguments.
\item
  This is just the same thing as I'm meaning by \textbf{structure}.
\item
  There is something about the Skippy example that's special; just the
  form of the argument shows that it is valid.
\end{itemize}
\end{frame}

\begin{frame}{Structure and Substitution}
\protect\hypertarget{structure-and-substitution}{}
When the structure of the argument does the guaranteeing, the guarantee
would persist through some substitutions.

\begin{description}
\tightlist
\item[~]
All kangaroos are wise.
\item[~]
Skippy is a kangaroo.
\item[\(\therefore\)]
Skippy is wise.
\end{description}
\end{frame}

\begin{frame}{Structure and Substitution}
\protect\hypertarget{structure-and-substitution-1}{}
When the structure of the argument does the guaranteeing, the guarantee
would persist through some substitutions.

\begin{description}
\tightlist
\item[~]
All kangaroos are wise.
\item[~]
Lucky is a kangaroo.
\item[\(\therefore\)]
Lucky is wise.
\end{description}
\end{frame}

\begin{frame}{Structure and Substitution}
\protect\hypertarget{structure-and-substitution-2}{}
When the structure of the argument does the guaranteeing, the guarantee
would persist through some substitutions.

\begin{description}
\tightlist
\item[~]
All koalas are wise.
\item[~]
Lucky is a koala.
\item[\(\therefore\)]
Lucky is wise.
\end{description}
\end{frame}

\begin{frame}{Structure and Substitution}
\protect\hypertarget{structure-and-substitution-3}{}
When the structure of the argument does the guaranteeing, the guarantee
would persist through some substitutions.

\begin{description}
\tightlist
\item[~]
All koalas are dishonest.
\item[~]
Lucky is a koala.
\item[\(\therefore\)]
Lucky is dishonest.
\end{description}
\end{frame}

\begin{frame}{Terminology}
\protect\hypertarget{terminology-1}{}
\begin{itemize}
\tightlist
\item
  We'll say an argument is \textbf{valid} just in case the truth of the
  premises guarantees the truth of the conclusion.
\item
  We'll say an argument is \textbf{formally valid}, or occasionally
  \textbf{logically valid}, if it is valid, and any argument with the
  same form/structure is valid.
\item
  We'll say an argument is \textbf{sound} just in case it is valid and
  has true premises.
\item
  We won't have much interest here in soundness; just in validity.
\end{itemize}
\end{frame}

\begin{frame}{Structure}
\protect\hypertarget{structure-1}{}
\begin{itemize}
\tightlist
\item
  You might have noticed that I never actually defined what it is for
  two arguments to share a structure.
\item
  I just said that changing a name (consistently throughout the
  argument), or changing a noun/adjective (consistently) preserves
  structure.
\item
  But I didn't say more than that.
\item
  Hold that thought! We'll have quite a lot to say as the course goes on
  about this notion of structure.
\end{itemize}
\end{frame}

\begin{frame}{Validity}
\protect\hypertarget{validity}{}
Here are some equivalent ways to define validity.

\begin{itemize}
\tightlist
\item
  An argument is valid if necessarily, when the premises are true, the
  conclusion is too.
\item
  An argument is valid if it is impossible for the premises to be true
  and the conclusion to not be true.
\end{itemize}

We'll use that last formulation a lot.
\end{frame}

\begin{frame}{For Next Time}
\protect\hypertarget{for-next-time}{}
We'll start our look at the most fundamental kind of logic:
propositional logic.
\end{frame}

\end{document}
