% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[BoldFont = SF Pro Rounded Semibold]{SF Pro Rounded}
  \setmathfont[]{STIX Two Math}
\fi
\usefonttheme{serif} % use mainfont rather than sansfont for slide text
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={305 Lecture 2.1 - Argument Forms},
  pdfauthor={Brian Weatherson},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\let\Tiny=\tiny

 \setbeamertemplate{navigation symbols}{} 

% \usetheme{Madrid}
 \usetheme[numbering=none, progressbar=foot]{metropolis}
 \usecolortheme{wolverine}
 \usepackage{color}
 \usepackage{MnSymbol}
% \usepackage{movie15}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\DeclareSymbolFont{symbolsC}{U}{txsyc}{m}{n}
\DeclareMathSymbol{\boxright}{\mathrel}{symbolsC}{128}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

 \usepackage{tikz-qtree}
% \usepackage{markdown}
%\RequirePackage{bussproofs}
\RequirePackage[tableaux]{prooftrees}
\usetikzlibrary{arrows.meta}
 \forestset{line numbering, close with = x}
% Allow for easy commas inside trees
\renewcommand{\,}{\text{, }}


\usepackage{tabulary}

\usepackage{open-logic-config}

\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\AtBeginSection[]
{
\begin{frame}
	\Huge{\color{darkblue} \insertsection}
\end{frame}
}

\renewenvironment*{quote}	
	{\list{}{\rightmargin   \leftmargin} \item } 	
	{\endlist }

\definecolor{darkgreen}{rgb}{0,0.7,0}
\definecolor{darkblue}{rgb}{0,0,0.8}

\newcommand{\starttab}{\begin{center}
\vspace{6pt}
\begin{tabular}}

\newcommand{\stoptab}{\end{tabular}
\vspace{6pt}
\end{center}
\noindent}


\newcommand{\sif}{\rightarrow}
\newcommand{\siff}{\leftrightarrow}
\newcommand{\EF}{\end{frame}}


\newcommand{\TreeStart}[1]{
%\end{frame}
\begin{frame}
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=north}}
%\Tree
}

\newcommand{\TreeEnd}{
\end{tikzpicture}
%\end{center}
}

\newcommand{\DisplayArg}[2]{
\begin{enumerate}
{#1}
\end{enumerate}
\vspace{-6pt}
\hrulefill

%\hspace{14pt} #2
%{\addtolength{\leftskip}{14pt} #2}
\begin{quote}
{\normalfont #2}
\end{quote}
\vspace{12pt}
}

\newenvironment{ProofTree}[1][1]{
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=south}}
}
{
\end{tikzpicture}
\end{center}
}

\newcommand{\TreeFrame}[2]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\begin{prooftree}{}
#1
\end{prooftree}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#2
%\end{markdown}
\end{columns}
}

\newcommand{\ScaledTreeFrame}[3]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\scalebox{#1}{
\begin{prooftree}{}
#2
\end{prooftree}
}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#3
%\end{markdown}
\end{columns}
}

\usepackage[bb=boondox]{mathalfa}
\DeclareMathAlphabet{\mathbx}{U}{BOONDOX-ds}{m}{n}
\SetMathAlphabet{\mathbx}{bold}{U}{BOONDOX-ds}{b}{n}
\DeclareMathAlphabet{\mathbbx} {U}{BOONDOX-ds}{b}{n}


\newenvironment{oltableau}{\center\tableau{}} %wff format={anchor = base west}}}
       {\endtableau\endcenter}
       
\newcommand{\formula}[1]{$#1$}

\usepackage{tabulary}
\usepackage{booktabs}

\def\begincols{\begin{columns}}
\def\begincol{\begin{column}}
\def\endcol{\end{column}}
\def\endcols{\end{columns}}

\usepackage[italic]{mathastext}
\usepackage{nicefrac}

\definecolor{mygreen}{RGB}{0, 100, 0}
\definecolor{mypink2}{RGB}{219, 48, 122}
\definecolor{dodgerblue}{RGB}{30,144,255}

%\def\True{\textcolor{dodgerblue}{\text{T}}}
%\def\False{\textcolor{red}{\text{F}}}

\def\True{\mathbb{T}}
\def\False{\mathbb{F}}

% This is because arguments didn't have enough space after them
\usepackage{etoolbox}
\AfterEndEnvironment{description}{\vspace{9pt}}
\AfterEndEnvironment{oltableau}{\vspace{9pt}}
\BeforeBeginEnvironment{oltableau}{\vspace{9pt}}
\AfterEndEnvironment{center}{\vspace{12pt}}
\BeforeBeginEnvironment{tabular}{\vspace{9pt}}

\setlength\heavyrulewidth{0pt}
\setlength\lightrulewidth{0pt}

%\def\toprule{}
%\def\bottomrule{}
%\def\midrule{}

\setbeamertemplate{caption}{\raggedright\insertcaption}

\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{305 Lecture 2.1 - Argument Forms}
\author{Brian Weatherson}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{Plan for This Lecture}
\protect\hypertarget{plan-for-this-lecture}{}
We're going to get a little clearer on the notions of `form' and
`structure' from last time.
\end{frame}

\begin{frame}{Associated Reading}
\protect\hypertarget{associated-reading}{}
forall x, chapter 4, ``First Steps to Symbolization''.
\end{frame}

\begin{frame}{Definition}
\protect\hypertarget{definition}{}
An argument is valid when

\begin{itemize}
\tightlist
\item
  The truth of the premises guarantees the truth of the conclusion.
\item
  Equivalently, when it is impossible for the premises to be true and
  the conclusion not true.
\end{itemize}
\end{frame}

\begin{frame}{Formal Validity}
\protect\hypertarget{formal-validity}{}
An argument is formally valid if

\begin{itemize}
\tightlist
\item
  It is valid; and
\item
  Any argument of the same form is valid.
\end{itemize}

In those cases, we infer that it is valid \textbf{because of} its form.
\end{frame}

\begin{frame}{Three Big Questions About Form}
\protect\hypertarget{three-big-questions-about-form}{}
\begin{enumerate}[<+->]
\tightlist
\item
  Which words get treated as part of the form, and which get replaced by
  symbols when converting into formal notation?
\item
  Which words have the same meaning and should be replaced by the same
  symbol, and which should be treated differently?
\item
  How do we break the sentence into its constituent parts?
\end{enumerate}
\end{frame}

\begin{frame}{Part of the Form}
\protect\hypertarget{part-of-the-form}{}
I'm going to have a fairly pluralist, pragmatist view on this question.

\begin{itemize}
\tightlist
\item
  This is not to say that pluralism or pragmatism is right - it's just
  that it's what I'll adopt. (It is to say that I think it's a
  reasonable approach.)
\item
  For some purposes the `formal words' are very few, like `and', `or',
  `not' etc.
\item
  For other purposes the set is more expansive.
\end{itemize}
\end{frame}

\begin{frame}{Part of the Form}
\protect\hypertarget{part-of-the-form-1}{}
\begin{itemize}
\tightlist
\item
  If you are studying the logic of causation, you might want to say that
  `caused' is a formal word.
\item
  And that would mean any two arguments that have this structure are of
  the same form.
\end{itemize}

\begin{description}
\tightlist
\item[~]
A caused B.
\item[~]
B caused C.
\item[\(\therefore\)]
A caused C.
\end{description}
\end{frame}

\begin{frame}{Part of the Form}
\protect\hypertarget{part-of-the-form-2}{}
\begin{itemize}
\tightlist
\item
  But for other purposes, `caused' is just another binary relation
  between events.
\item
  So those arguments will have the same form as these ones.
\end{itemize}

\begin{description}
\tightlist
\item[~]
A happened near B.
\item[~]
B happened near C.
\item[\(\therefore\)]
A happened near C.
\end{description}
\end{frame}

\begin{frame}{Part of the Form}
\protect\hypertarget{part-of-the-form-3}{}
\begin{itemize}
\tightlist
\item
  What I mean by having a `pluralist, pragmatist' view about form is
  that there is no objectively correct answer about whether these are
  the same form or not.
\item
  For some purposes it will be helpful to treat them as having the same
  form.
\item
  For other purposes it will not be.
\item
  We're going to start with a notion of form where lots and lots of
  things have the same form, so very few arguments are valid in virtue
  of form.
\end{itemize}
\end{frame}

\begin{frame}{Three Big Questions About Form}
\protect\hypertarget{three-big-questions-about-form-1}{}
\begin{enumerate}
\tightlist
\item
  Which words get treated as part of the form, and which get replaced by
  symbols when converting into formal notation?
\item
  \textbf{Which words have the same meaning and should be replaced by
  the same symbol, and which should be treated differently?}
\item
  How do we break the sentence into its constituent parts?
\end{enumerate}
\end{frame}

\begin{frame}{Same/Different}
\protect\hypertarget{samedifferent}{}
There are problems in both directions here.

\begin{itemize}
\tightlist
\item
  Sometimes we want to treat words with different spelling as
  representing the same meaning - e.g., `kangaroo' and `kangaroos'.
\item
  And sometimes we want to treat words with the same spelling as
  expressing different meanings.
\end{itemize}
\end{frame}

\begin{frame}{Relating the First Two Questions}
\protect\hypertarget{relating-the-first-two-questions}{}
Some kinds of ambiguity are `resolved' by treating the word as part of
the formal system. Consider two possible interpretations of this
sentence.

\begin{quote}
Nothing is better than world peace
\end{quote}
\end{frame}

\begin{frame}{``Nothing is better than world peace''}
\protect\hypertarget{nothing-is-better-than-world-peace}{}
\begin{enumerate}[<+->]
\tightlist
\item
  There is no thing such that it is better than world peace.
\item
  It is better that there is nothing than that there is world peace.
\end{enumerate}
\end{frame}

\begin{frame}{Ambiguity and Form}
\protect\hypertarget{ambiguity-and-form}{}
\begin{itemize}
\tightlist
\item
  If you include words like `nothing' as formal words, then the two
  readings of the sentence will have different formal representations.
\item
  And the logic called `predicate logic' does exactly that; it treats
  `nothing' as one of the formal words.
\item
  We're not covering that in this course, though you have have covered
  it elsewhere.
\item
  Some more expansive accounts of what should be a formal word are moved
  by the thought that treating more words as formal removes some
  possible ambiguities.
\end{itemize}
\end{frame}

\begin{frame}{Three Big Questions About Form}
\protect\hypertarget{three-big-questions-about-form-2}{}
\begin{enumerate}
\tightlist
\item
  Which words get treated as part of the form, and which get replaced by
  symbols when converting into formal notation?
\item
  Which words have the same meaning and should be replaced by the same
  symbol, and which should be treated differently?
\item
  \textbf{How do we break the sentence into its constituent parts?}
\end{enumerate}
\end{frame}

\begin{frame}{Three Formal Tools}
\protect\hypertarget{three-formal-tools}{}
\begin{enumerate}[<+->]
\tightlist
\item
  Replacing words/sentences with letters.
\item
  Using parentheses.
\item
  Replacing `formal words' with symbols.
\end{enumerate}
\end{frame}

\begin{frame}{For Next Time}
\protect\hypertarget{for-next-time}{}
We'll look at how propositional logic carries out these three tasks.
\end{frame}

\end{document}
