% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[BoldFont = SF Pro Rounded Semibold]{SF Pro Rounded}
  \setmathfont[]{STIX Two Math}
\fi
\usefonttheme{serif} % use mainfont rather than sansfont for slide text
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={305 Lecture 3.5 - Introducing Truth Trees},
  pdfauthor={Brian Weatherson},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\let\Tiny=\tiny

 \setbeamertemplate{navigation symbols}{} 

% \usetheme{Madrid}
 \usetheme[numbering=none, progressbar=foot]{metropolis}
 \usecolortheme{wolverine}
 \usepackage{color}
 \usepackage{MnSymbol}
% \usepackage{movie15}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\DeclareSymbolFont{symbolsC}{U}{txsyc}{m}{n}
\DeclareMathSymbol{\boxright}{\mathrel}{symbolsC}{128}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

 \usepackage{tikz-qtree}
% \usepackage{markdown}
%\RequirePackage{bussproofs}
\RequirePackage[tableaux]{prooftrees}
\usetikzlibrary{arrows.meta}
 \forestset{line numbering, close with = x}
% Allow for easy commas inside trees
\renewcommand{\,}{\text{, }}


\usepackage{tabulary}

\usepackage{open-logic-config}

\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\AtBeginSection[]
{
\begin{frame}
	\Huge{\color{darkblue} \insertsection}
\end{frame}
}

\renewenvironment*{quote}	
	{\list{}{\rightmargin   \leftmargin} \item } 	
	{\endlist }

\definecolor{darkgreen}{rgb}{0,0.7,0}
\definecolor{darkblue}{rgb}{0,0,0.8}

\newcommand{\starttab}{\begin{center}
\vspace{6pt}
\begin{tabular}}

\newcommand{\stoptab}{\end{tabular}
\vspace{6pt}
\end{center}
\noindent}


\newcommand{\sif}{\rightarrow}
\newcommand{\siff}{\leftrightarrow}
\newcommand{\EF}{\end{frame}}


\newcommand{\TreeStart}[1]{
%\end{frame}
\begin{frame}
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=north}}
%\Tree
}

\newcommand{\TreeEnd}{
\end{tikzpicture}
%\end{center}
}

\newcommand{\DisplayArg}[2]{
\begin{enumerate}
{#1}
\end{enumerate}
\vspace{-6pt}
\hrulefill

%\hspace{14pt} #2
%{\addtolength{\leftskip}{14pt} #2}
\begin{quote}
{\normalfont #2}
\end{quote}
\vspace{12pt}
}

\newenvironment{ProofTree}[1][1]{
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=south}}
}
{
\end{tikzpicture}
\end{center}
}

\newcommand{\TreeFrame}[2]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\begin{prooftree}{}
#1
\end{prooftree}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#2
%\end{markdown}
\end{columns}
}

\newcommand{\ScaledTreeFrame}[3]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\scalebox{#1}{
\begin{prooftree}{}
#2
\end{prooftree}
}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#3
%\end{markdown}
\end{columns}
}

\usepackage[bb=boondox]{mathalfa}
\DeclareMathAlphabet{\mathbx}{U}{BOONDOX-ds}{m}{n}
\SetMathAlphabet{\mathbx}{bold}{U}{BOONDOX-ds}{b}{n}
\DeclareMathAlphabet{\mathbbx} {U}{BOONDOX-ds}{b}{n}


\newenvironment{oltableau}{\center\tableau{}} %wff format={anchor = base west}}}
       {\endtableau\endcenter}
       
\newcommand{\formula}[1]{$#1$}

\usepackage{tabulary}
\usepackage{booktabs}

\def\begincols{\begin{columns}}
\def\begincol{\begin{column}}
\def\endcol{\end{column}}
\def\endcols{\end{columns}}

\usepackage[italic]{mathastext}
\usepackage{nicefrac}

\definecolor{mygreen}{RGB}{0, 100, 0}
\definecolor{mypink2}{RGB}{219, 48, 122}
\definecolor{dodgerblue}{RGB}{30,144,255}

%\def\True{\textcolor{dodgerblue}{\text{T}}}
%\def\False{\textcolor{red}{\text{F}}}

\def\True{\mathbb{T}}
\def\False{\mathbb{F}}

% This is because arguments didn't have enough space after them
\usepackage{etoolbox}
\AfterEndEnvironment{description}{\vspace{9pt}}
\AfterEndEnvironment{oltableau}{\vspace{9pt}}
\BeforeBeginEnvironment{oltableau}{\vspace{9pt}}
\AfterEndEnvironment{center}{\vspace{12pt}}
\BeforeBeginEnvironment{tabular}{\vspace{9pt}}

\setlength\heavyrulewidth{0pt}
\setlength\lightrulewidth{0pt}

%\def\toprule{}
%\def\bottomrule{}
%\def\midrule{}

\setbeamertemplate{caption}{\raggedright\insertcaption}

\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{305 Lecture 3.5 - Introducing Truth Trees}
\author{Brian Weatherson}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{Plan}
\protect\hypertarget{plan}{}
This lecture introduces a new way of testing for validity - truth trees.
\end{frame}

\begin{frame}{Associated Reading}
\protect\hypertarget{associated-reading}{}
Boxes and Diamonds, section 2.1
\end{frame}

\begin{frame}{What Tableaux Are}
\protect\hypertarget{what-tableaux-are}{}
\begin{itemize}
\tightlist
\item
  A way for determining whether some combinations are logically
  possible.
\item
  That can be used for determining whether some arguments are valid - if
  the truth of the premises and the falsity of the conclusion is not
  logically possible; then the argument is valid.
\end{itemize}
\end{frame}

\begin{frame}{Structure}
\protect\hypertarget{structure}{}
\begin{itemize}
\tightlist
\item
  Start out writing the things you care about.\pause
\item
  Each time one of those things implies that some other things must be
  the case, write those down too.
\item
  For example, if you write down that \(A \wedge B\) is true, also write
  down that \(A\) is true and that \(B\) is true. \pause
\item
  Each time there are multiple ways to make something you've written
  true, create multiple branches for those ways.
\item
  For example, if you write down that \(A \vee B\) is true, create a
  branch where \(A\) is true, and a branch where \(B\) is true.
\end{itemize}
\end{frame}

\begin{frame}{Closing}
\protect\hypertarget{closing}{}
\begin{itemize}
\tightlist
\item
  A branch of a tableau is closed if it contains that some particular
  claim has incompatible truth values.
\item
  For now, this means that one sentence is both true and false.\pause
\item
  The whole tableau is closed if every branch is closed.
\end{itemize}
\end{frame}

\begin{frame}{What Closure Means}
\protect\hypertarget{what-closure-means}{}
\begin{itemize}
\tightlist
\item
  If the tableau is closed, then the initial assumptions cannot be true
  together.
\item
  If you are evaluating an argument, this means that the argument is
  valid.\pause
\item
  If you start the tableau by just saying that one sentence is false,
  the closure of the tableau means that that sentence is a logical
  truth.
\end{itemize}
\end{frame}

\begin{frame}{Benefits of Tableaux}
\protect\hypertarget{benefits-of-tableaux}{}
Tableaux have two big benefits over truth tables.

\begin{enumerate}
\tightlist
\item
  They don't grow exponentially when you increase the number of
  variables.\pause
\item
  They can be generalised to things beyond propositional logic.
\end{enumerate}

We are introducing them here because of point 2.
\end{frame}

\begin{frame}{Open Tableaux}
\protect\hypertarget{open-tableaux}{}
Here is something the book doesn't make a big deal of, but is kind of
important.

\begin{itemize}
\tightlist
\item
  A closed tableau can show that an argument is valid.
\item
  An open \textbf{and completed} tableau can show that an argument is
  invalid.\pause
\item
  The trick here is that it's hard to tell when a tableau is completed
  in the relevant sense.
\item
  This will be easier to illustrate in practice than in theory, so let's
  start building tableau up.
\end{itemize}
\end{frame}

\begin{frame}{Signs}
\protect\hypertarget{signs}{}
The system we are using is what is called a \textbf{signed tableau}
system.

\begin{itemize}
\tightlist
\item
  That means that every line consists of two parts.
\item
  The bigger, second, part is a sentence.
\item
  The first part is a \textbf{sign}, which for now is a truth value.
\item
  That is, it is either \(\True\) or \(\False\).
\end{itemize}
\end{frame}

\begin{frame}{What Lines Mean}
\protect\hypertarget{what-lines-mean}{}
So each line either says that a particular sentence is true, or says
that it is false.

\begin{itemize}
\tightlist
\item
  The book for some reason includes the word `might' here.
\item
  That's misleading; what they should say is that each line says what is
  true given (a) the starting assumptions and (b) the assumptions we
  made for branching purposes.
\end{itemize}
\end{frame}

\begin{frame}{Numbering And Annotation}
\protect\hypertarget{numbering-and-annotation}{}
The system we're using also has two other distinctive features.

\begin{enumerate}
\tightlist
\item
  We number each of the lines. \pause
\item
  We list why we're writing down each line to the right of the tree.
  \pause
\end{enumerate}

Both of these are somewhat idiosyncratic, though not abnormal. Unlike
the truth tables, there just aren't well defined conventions for how to
write these things out.
\end{frame}

\begin{frame}{For Next Time}
\protect\hypertarget{for-next-time}{}
We will start looking at the rules that get used for building truth
trees.
\end{frame}

\end{document}
