% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[BoldFont = SF Pro Rounded Semibold]{SF Pro Rounded}
  \setmathfont[]{STIX Two Math}
\fi
\usefonttheme{serif} % use mainfont rather than sansfont for slide text
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={305 Lecture 7.2 - About Probability Functions},
  pdfauthor={Brian Weatherson},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
\usepackage{caption}
% Make caption package work with longtable
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\let\Tiny=\tiny

 \setbeamertemplate{navigation symbols}{} 

% \usetheme{Madrid}
 \usetheme[numbering=none, progressbar=foot]{metropolis}
 \usecolortheme{wolverine}
 \usepackage{color}
 \usepackage{MnSymbol}
% \usepackage{movie15}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\DeclareSymbolFont{symbolsC}{U}{txsyc}{m}{n}
\DeclareMathSymbol{\boxright}{\mathrel}{symbolsC}{128}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

 \usepackage{tikz-qtree}
% \usepackage{markdown}
%\RequirePackage{bussproofs}
\RequirePackage[tableaux]{prooftrees}
\usetikzlibrary{arrows.meta}
 \forestset{line numbering, close with = x}
% Allow for easy commas inside trees
\renewcommand{\,}{\text{, }}


\usepackage{tabulary}

\usepackage{open-logic-config}

\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\AtBeginSection[]
{
\begin{frame}
	\Huge{\color{darkblue} \insertsection}
\end{frame}
}

\renewenvironment*{quote}	
	{\list{}{\rightmargin   \leftmargin} \item } 	
	{\endlist }

\definecolor{darkgreen}{rgb}{0,0.7,0}
\definecolor{darkblue}{rgb}{0,0,0.8}

\newcommand{\starttab}{\begin{center}
\vspace{6pt}
\begin{tabular}}

\newcommand{\stoptab}{\end{tabular}
\vspace{6pt}
\end{center}
\noindent}


\newcommand{\sif}{\rightarrow}
\newcommand{\siff}{\leftrightarrow}
\newcommand{\EF}{\end{frame}}


\newcommand{\TreeStart}[1]{
%\end{frame}
\begin{frame}
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=north}}
%\Tree
}

\newcommand{\TreeEnd}{
\end{tikzpicture}
%\end{center}
}

\newcommand{\DisplayArg}[2]{
\begin{enumerate}
{#1}
\end{enumerate}
\vspace{-6pt}
\hrulefill

%\hspace{14pt} #2
%{\addtolength{\leftskip}{14pt} #2}
\begin{quote}
{\normalfont #2}
\end{quote}
\vspace{12pt}
}

\newenvironment{ProofTree}[1][1]{
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=south}}
}
{
\end{tikzpicture}
\end{center}
}

\newcommand{\TreeFrame}[2]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\begin{prooftree}{}
#1
\end{prooftree}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#2
%\end{markdown}
\end{columns}
}

\newcommand{\ScaledTreeFrame}[3]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\scalebox{#1}{
\begin{prooftree}{}
#2
\end{prooftree}
}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#3
%\end{markdown}
\end{columns}
}

\usepackage[bb=boondox]{mathalfa}
\DeclareMathAlphabet{\mathbx}{U}{BOONDOX-ds}{m}{n}
\SetMathAlphabet{\mathbx}{bold}{U}{BOONDOX-ds}{b}{n}
\DeclareMathAlphabet{\mathbbx} {U}{BOONDOX-ds}{b}{n}


\newenvironment{oltableau}{\center\tableau{}} %wff format={anchor = base west}}}
       {\endtableau\endcenter}
       
\newcommand{\formula}[1]{$#1$}

\usepackage{tabulary}
\usepackage{booktabs}

\def\begincols{\begin{columns}}
\def\begincol{\begin{column}}
\def\endcol{\end{column}}
\def\endcols{\end{columns}}

\usepackage[italic]{mathastext}
\usepackage{nicefrac}

\definecolor{mygreen}{RGB}{0, 100, 0}
\definecolor{mypink2}{RGB}{219, 48, 122}
\definecolor{dodgerblue}{RGB}{30,144,255}

%\def\True{\textcolor{dodgerblue}{\text{T}}}
%\def\False{\textcolor{red}{\text{F}}}

\def\True{\mathbb{T}}
\def\False{\mathbb{F}}

% This is because arguments didn't have enough space after them
\usepackage{etoolbox}
\AfterEndEnvironment{description}{\vspace{9pt}}
\AfterEndEnvironment{oltableau}{\vspace{9pt}}
\BeforeBeginEnvironment{oltableau}{\vspace{9pt}}
\AfterEndEnvironment{center}{\vspace{12pt}}
\BeforeBeginEnvironment{tabular}{\vspace{9pt}}

\setlength\heavyrulewidth{0pt}
\setlength\lightrulewidth{0pt}

%\def\toprule{}
%\def\bottomrule{}
%\def\midrule{}

\setbeamertemplate{caption}{\raggedright\insertcaption}

\setlength\lightrulewidth{0.3pt}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{305 Lecture 7.2 - About Probability Functions}
\author{Brian Weatherson}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{Plan}
\protect\hypertarget{plan}{}
\begin{itemize}
\tightlist
\item
  This lecture looks at some general features of probability functions,
  and looks at some ways to think probabilistically about real world
  events.
\end{itemize}
\end{frame}

\begin{frame}{Associated Reading}
\protect\hypertarget{associated-reading}{}
Odds and Ends, Chapter 5
\end{frame}

\begin{frame}{Scale}
\protect\hypertarget{scale}{}
\[
0 \leq \Pr(A) \leq 1
\]
\end{frame}

\begin{frame}{Negation}
\protect\hypertarget{negation}{}
\[
\Pr(\neg A) = 1 - \Pr(A)
\]
\end{frame}

\begin{frame}{Excluded Middle}
\protect\hypertarget{excluded-middle}{}
\[
\Pr(A) + \Pr(\neg A) = 1
\]
\end{frame}

\begin{frame}{Partition}
\protect\hypertarget{partition}{}
Some events \(A_1, \dots A_n\) form a partition if, necessarily, exactly
one of them is true.

\begin{itemize}
\tightlist
\item
  So they are \textbf{exclusive} - you can't have any two of them both
  be true.
\item
  And they are \textbf{exhaustive} - you have to have at least one true.
\end{itemize}
\end{frame}

\begin{frame}{Partition}
\protect\hypertarget{partition-1}{}
If \(A_1, \dots A_n\) form a partition then

\[
\Pr(A_1) + \dots + \Pr(A_n) = 1
\]
\end{frame}

\begin{frame}{Exclusive}
\protect\hypertarget{exclusive}{}
If \(A, B\) are exclusive

\[
\Pr(A \vee B) = \Pr(A) + \Pr(B)
\]
\end{frame}

\begin{frame}{General Principle}
\protect\hypertarget{general-principle}{}
\[
\Pr(A) + \Pr(B) = \Pr(A \vee B) + \Pr(A \wedge B)
\]

\pause

It's worth thinking through why this is true in terms of possibilities.
\end{frame}

\begin{frame}{Trees}
\protect\hypertarget{trees}{}
\begin{itemize}
\tightlist
\item
  Often, we can't just write down numbers for the possibilities.
\item
  But we can write down, or at least make reasonable guesses about,
  numbers for certain events.
\item
  And we can think about how things are likely to go given those events
  happen.
\item
  This is the tree structure that is used in \emph{Odds and Ends}.
\end{itemize}
\end{frame}

\begin{frame}{Academic Forecasting}
\protect\hypertarget{academic-forecasting}{}
\begin{itemize}
\tightlist
\item
  So let's say you're interested in the probability that we have in
  person classes all the way through Fall semester.
\item
  The vaccine rollout is going much better than I ever expected.
\item
  But we saw how bad things got last Thanksgiving and Christmas, and you
  can see why that might make you worried.
\item
  But it's hard to put into numbers how it affects things.
\end{itemize}
\end{frame}

\begin{frame}{One Method}
\protect\hypertarget{one-method}{}
\begin{itemize}
\tightlist
\item
  Divide up the state space.
\item
  Work out the probability of being in one or other part of the space.
\item
  Work out the probability of classes running all semester in each part
  of the space.
\item
  Add things up.
\end{itemize}
\end{frame}

\begin{frame}{Nothing is Reliable}
\protect\hypertarget{nothing-is-reliable}{}
\begin{itemize}[<+->]
\tightlist
\item
  There are a lot of ways to do this.
\item
  You could divide up the space by how many cases there are in Michigan.
\item
  Or you could divide up the space by how many other universities shut
  down.
\item
  Or, and let's work with this one, you could divide it up by how many
  people are vaccinated by the the start of term.
\end{itemize}
\end{frame}

\begin{frame}{Three States}
\protect\hypertarget{three-states}{}
\begin{enumerate}
\tightlist
\item
  Only high risk staff and students are vaccinated.
\item
  More than just the high risk folks are vaccinated, but there are still
  shortages.
\item
  Vaccine for all.
\end{enumerate}
\end{frame}

\begin{frame}{Two Step Process}
\protect\hypertarget{two-step-process}{}
\begin{enumerate}
\tightlist
\item
  Work out (or at least estimate) probability of each state.
\item
  Work out probability of the university staying open within each of
  those states.
\end{enumerate}

This will involve a lot of guesswork - do not make investment decisions
on the basis of the numbers I'm about to use because they are really
pulled out of the air - but it's a very helpful heuristic to at least
approximate the reality.
\end{frame}

\begin{frame}{Probabilities of States}
\protect\hypertarget{probabilities-of-states}{}
This could be embarrassingly out of date by the time we even do the
classes, but let's say the probabilities look like this.

\begin{itemize}
\tightlist
\item
  High risk only - 20\%, or 0.2.
\item
  Limited quantities of vaccine - 50\%, or 0.5.
\item
  Unlimited quantities of vaccine - 30\%, or 0.3.
\end{itemize}
\end{frame}

\begin{frame}{Class Probabilities}
\protect\hypertarget{class-probabilities}{}
Then we want to work out how likely it is that we get a full semester of
classes in given each of these three possibilities.
\end{frame}

\begin{frame}{High Risk Only}
\protect\hypertarget{high-risk-only}{}
Given how badly things went last Fall, I'm not sure what we'd do in this
situation. Maybe we'd muddle through and it would work. But I suspect
that it's most likely we'd be planning to shut down at Thanksgiving, so
not strictly a full semester.

\begin{itemize}[<+->]
\tightlist
\item
  Let's call it a 30\% chance of staying open all semester in that
  situation.
\item
  I really, and I can't stress this enough, want to emphasise that I'm
  making these numbers up for the purposes of an in class example.
\end{itemize}
\end{frame}

\begin{frame}{Limited Vaccine}
\protect\hypertarget{limited-vaccine}{}
This starts to feel more like something where we can get through the
term. For one thing, we might just tought it out even if there is a bit
of an outbreak. For another, I said limited in September; maybe by
October/November it would be unlimited. So it feels pretty likely.

\begin{itemize}
\tightlist
\item
  Let's say there is an 80\% chance we make it all the way through
  semester in that situation, and only 20\% that we don't.
\end{itemize}
\end{frame}

\begin{frame}{Unlimited vaccine}
\protect\hypertarget{unlimited-vaccine}{}
Now it's hard to see why the university would shut down. Rather than
fuss about the details, let's just simplify and say that in this (very
unlikely) scenario, it's 100\% likely that we'll get all classes in.
\end{frame}

\begin{frame}{The Giant Table}
\protect\hypertarget{the-giant-table}{}
\begin{longtable}[]{@{}lcc@{}}
\toprule
& Full Semester & Some shutdown \\ \addlinespace
\midrule
\endhead
No vaccine & \(0.2 \times 0.3 = 0.06\) &
\(0.2 \times 0.7 = 0.14\) \\ \addlinespace
Limited vaccine & \(0.5 \times 0.8 = 0.40\) &
\(0.5 \times 0.2 = 0.10\) \\ \addlinespace
Unlimited vaccine & \(0.3 \times 1.0 = 0.30\) &
\(0.3 \times 0.0 = 0.00\) \\ \addlinespace
\bottomrule
\end{longtable}

So the probability that we get a full semester in (given these literally
made up assumptions) is \(0.06 + 0.40 + 0.30 = 0.76\), or 76\%.
\end{frame}

\begin{frame}{Probabilities and Errors}
\protect\hypertarget{probabilities-and-errors}{}
\begin{itemize}
\tightlist
\item
  The error bars on that calculation are massive.
\item
  But it's a kind of sanity check on how you think things are going.
\item
  Especially in situations where only a handful of paths lead to a
  salient result (like in playoffs in sports, or when thinking about the
  likelihood of a particular challenger winning), doing the tree even
  with favorable numbers can give you a conservative estimate of some
  probability.
\end{itemize}
\end{frame}

\begin{frame}{Three Cases Where This is More Precise}
\protect\hypertarget{three-cases-where-this-is-more-precise}{}
\begin{enumerate}
\tightlist
\item
  Probabilities are stipulated
\item
  Probabilities are due to well understood chance processes (like
  gambling devices)
\item
  Probabilities are derived from very large data sets
\end{enumerate}
\end{frame}

\begin{frame}{For Next Time}
\protect\hypertarget{for-next-time}{}
\begin{itemize}
\tightlist
\item
  We will look at an example where the probabilities are indeed
  stipulated.
\end{itemize}
\end{frame}

\end{document}
