% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[BoldFont = SF Pro Rounded Semibold]{SF Pro Rounded}
  \setmathfont[]{STIX Two Math}
\fi
\usefonttheme{serif} % use mainfont rather than sansfont for slide text
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={305 Lecture 8.2 - Updating on Multiple Data Points},
  pdfauthor={Brian Weatherson},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\let\Tiny=\tiny

 \setbeamertemplate{navigation symbols}{} 

% \usetheme{Madrid}
 \usetheme[numbering=none, progressbar=foot]{metropolis}
 \usecolortheme{wolverine}
 \usepackage{color}
 \usepackage{MnSymbol}
% \usepackage{movie15}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\DeclareSymbolFont{symbolsC}{U}{txsyc}{m}{n}
\DeclareMathSymbol{\boxright}{\mathrel}{symbolsC}{128}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\usepackage{tikz-qtree}
% \usepackage{markdown}
%\RequirePackage{bussproofs}
\usetikzlibrary{arrows.meta}
\RequirePackage[tableaux]{prooftrees}
\forestset{line numbering, close with = x}
% Allow for easy commas inside trees
\renewcommand{\,}{\text{, }}


\usepackage{tabulary}

\usepackage{open-logic-config}

\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\AtBeginSection[]
{
\begin{frame}
	\Huge{\color{darkblue} \insertsection}
\end{frame}
}

\renewenvironment*{quote}	
	{\list{}{\rightmargin   \leftmargin} \item } 	
	{\endlist }

\definecolor{darkgreen}{rgb}{0,0.7,0}
\definecolor{darkblue}{rgb}{0,0,0.8}

\newcommand{\starttab}{\begin{center}
\vspace{6pt}
\begin{tabular}}

\newcommand{\stoptab}{\end{tabular}
\vspace{6pt}
\end{center}
\noindent}


\newcommand{\sif}{\rightarrow}
\newcommand{\siff}{\leftrightarrow}
\newcommand{\EF}{\end{frame}}


\newcommand{\TreeStart}[1]{
%\end{frame}
\begin{frame}
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=north}}
%\Tree
}

\newcommand{\TreeEnd}{
\end{tikzpicture}
%\end{center}
}

\newcommand{\DisplayArg}[2]{
\begin{enumerate}
{#1}
\end{enumerate}
\vspace{-6pt}
\hrulefill

%\hspace{14pt} #2
%{\addtolength{\leftskip}{14pt} #2}
\begin{quote}
{\normalfont #2}
\end{quote}
\vspace{12pt}
}

\newenvironment{ProofTree}[1][1]{
\begin{center}
\begin{tikzpicture}[scale=#1]
\tikzset{every tree node/.style={align=center,anchor=south}}
}
{
\end{tikzpicture}
\end{center}
}

\newcommand{\TreeFrame}[2]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\begin{prooftree}{}
#1
\end{prooftree}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#2
%\end{markdown}
\end{columns}
}

\newcommand{\ScaledTreeFrame}[3]{
\begin{columns}[c]
\column{0.5\textwidth}
\begin{center}
\scalebox{#1}{
\begin{prooftree}{}
#2
\end{prooftree}
}
\end{center}
\column{0.45\textwidth}
%\begin{markdown}
#3
%\end{markdown}
\end{columns}
}

\usepackage[bb=boondox]{mathalfa}
\DeclareMathAlphabet{\mathbx}{U}{BOONDOX-ds}{m}{n}
\SetMathAlphabet{\mathbx}{bold}{U}{BOONDOX-ds}{b}{n}
\DeclareMathAlphabet{\mathbbx} {U}{BOONDOX-ds}{b}{n}


\newenvironment{oltableau}{\center\tableau{}} %wff format={anchor = base west}}}
       {\endtableau\endcenter}
       
\newcommand{\formula}[1]{$#1$}

\usepackage{tabulary}
\usepackage{booktabs}

\def\begincols{\begin{columns}}
\def\begincol{\begin{column}}
\def\endcol{\end{column}}
\def\endcols{\end{columns}}

\usepackage[italic]{mathastext}
\usepackage{nicefrac}

\definecolor{mygreen}{RGB}{0, 100, 0}
\definecolor{mypink2}{RGB}{219, 48, 122}
\definecolor{dodgerblue}{RGB}{30,144,255}

%\def\True{\textcolor{dodgerblue}{\text{T}}}
%\def\False{\textcolor{red}{\text{F}}}

\def\True{\mathbb{T}}
\def\False{\mathbb{F}}

% This is because arguments didn't have enough space after them
\usepackage{etoolbox}
\AfterEndEnvironment{description}{\vspace{9pt}}
\AfterEndEnvironment{oltableau}{\vspace{9pt}}
\BeforeBeginEnvironment{oltableau}{\vspace{9pt}}
\AfterEndEnvironment{center}{\vspace{12pt}}
\BeforeBeginEnvironment{tabular}{\vspace{9pt}}

\setlength\heavyrulewidth{0pt}
\setlength\lightrulewidth{0pt}

%\def\toprule{}
%\def\bottomrule{}
%\def\midrule{}

\setbeamertemplate{caption}{\raggedright\insertcaption}

\setlength\lightrulewidth{0.3pt}
\AfterEndEnvironment{center}{\vspace{-3pt}}
\AfterEndEnvironment{longtable}{\vspace{-6pt}}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{305 Lecture 8.2 - Updating on Multiple Data Points}
\author{Brian Weatherson}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{Plan}
\protect\hypertarget{plan}{}
\begin{itemize}
\tightlist
\item
  To look at what happens when we learn two facts about some
  environment.
\end{itemize}
\end{frame}

\begin{frame}{Associated Reading}
\protect\hypertarget{associated-reading}{}
Odds and Ends, Chapter 9
\end{frame}

\begin{frame}{Conditional Independence}
\protect\hypertarget{conditional-independence}{}
In a lot of cases, the two data points we get will not be
probabilistically independent, but they will be \textbf{conditionally
independent}.

That is, if \(B_1\) and \(B_2\) are the data points, and \(X\) is an
arbitrary hypothesis (like \(A, \neg A\)), we will have

\begin{quote}
\(\Pr(B_1 | X)\Pr(B_2 | X) = \Pr(B_1 \wedge B_2 | X)\)
\end{quote}
\end{frame}

\begin{frame}{Biased Coins}
\protect\hypertarget{biased-coins}{}
Here is one kind of case where the happens.

\begin{itemize}
\tightlist
\item
  We have a bunch of biased coins. For each of them, there is a
  probability \(p\) of heads on an arbitrary flip, but we don't know
  what that is. \pause
\item
  The results of two flips of the same coin are not indepdendent.
\item
  If one flip lands heads, that is evidence of a bias towards heads, and
  hence it increases the probability of heads on the next flip. \pause
\item
  But conditional on a hypothesis about the bias of the coin, the flips
  are independent.
\end{itemize}
\end{frame}

\begin{frame}{Skilled Activity}
\protect\hypertarget{skilled-activity}{}
A perhaps more real-life case of this is skilled action, like shooting
free throws.

\begin{itemize}
\tightlist
\item
  The success of one attempt is not independent of the success of the
  previous.
\item
  But conditional on the skill of the actor, the attempts are (probably,
  more or less) independent.
\end{itemize}
\end{frame}

\begin{frame}{Sampling With Replacement}
\protect\hypertarget{sampling-with-replacement}{}
Drawing from a selection \textbf{with replacement} produces conditional
independence.

\begin{itemize}
\tightlist
\item
  If I don't know how many black marbles are in an urn, then drawing a
  black marble \textbf{and replacing it} will be evidence that the next
  marble is black.
\item
  But conditional on a hypothesis about the nature of the urn, the draws
  with replacement will be independent.
\end{itemize}
\end{frame}

\begin{frame}{Yesterday, Today, Tomorrow}
\protect\hypertarget{yesterday-today-tomorrow}{}
This is a little off topic, but a lot of real world phenomena satisfy
(roughly) the following condition.

\begin{itemize}
\tightlist
\item
  How things were yesterday is a good (probabilistic) guide to how
  things will be tomorrow.
\item
  So how things will be tomorrow is not independent of how things were
  yesterday. \pause
\item
  But, conditional on how things are today, how things were yesterday
  and will be tomorrow are independent.
\item
  Knowing how things were yesterday doesn't tell you any more about how
  things will be tomorrow once you know how things are today.
\end{itemize}
\end{frame}

\begin{frame}{Markov Chains}
\protect\hypertarget{markov-chains}{}
A chain of events where every event is probabilistically dependent on
the previous one, but only on the previous one, is called a
\textbf{Markov Chain}. \pause

\begin{itemize}
\tightlist
\item
  Lots of real world processes are (more or less) Markov Chains.
\item
  Weather systems, for instance, are probably more or less Markov
  Chains.
\item
  And lots of ecological models assume that animal populations are
  Markov Chains.
\item
  And the core idea is just conditional independence.
\end{itemize}
\end{frame}

\begin{frame}{Conditional Independence}
\protect\hypertarget{conditional-independence-1}{}
In cases where the data points \(B_1\) and \(B_2\) are independent, we
have an easy story about how to work out the probabilities.

\begin{quote}
\(\Pr(B_1 \wedge B_2 | X) = \Pr(B_1 | X)\Pr(B_2 | X)\)
\end{quote}
\end{frame}

\begin{frame}{Same Event}
\protect\hypertarget{same-event}{}
There is an even simpler formula where \(B_1\) and \(B_2\) are the
`same' event, like the coin landing heads both time, or the same color
marble being drawn.

\begin{quote}
\(\Pr(B_1 \wedge B_2 | X) = \Pr(B_1 | X)^2\)
\end{quote}
\end{frame}

\begin{frame}{For Next Time}
\protect\hypertarget{for-next-time}{}
We'll start illustrating this with some worked examples.
\end{frame}

\end{document}
